---
title: "bias variance tradeoff regression"
author: "Ryan Attwood"
date: "`r Sys.Date()`"
output: html_document
---
# test set


set.seed(2)
sim_mat <- mvtnorm::rmvnorm(nrow(data_for_var), sigma = cor(data_for_var))
sim_unif <- pnorm(sim_mat)
colnames(sim_unif)<-colnames(data_for_var)


sim<-data_for_var

sim$percent_black_2010 <- qgamma(sim_unif[,"percent_black_2010"], shape=0.4402158, rate = 0.05254412 )
sim$Direct_farm_sales_per_capita_2012 <- qgamma(sim_unif[,"Direct_farm_sales_per_capita_2012"], shape=0.5355279 , rate = 0.06477432 )
sim$percent_white_2010<-truncnorm::qtruncnorm(sim_unif[,"percent_white_2010"], a=0, b=100, mean=160, sd = 45)
sim$percent_pacific_islander_2010<-qgamma(sim_unif[,"percent_pacific_islander_2010"], shape=0.2740278  , rate = 4.224056)
sim$CSA_2013<-qgamma(sim_unif[,"CSA_2013"], shape=0.3002073  , rate = 2465.577  )

st_fit <- sstdFit(data_for_var$percent_native_american_2010)
sim$percent_native_american_2010 <- 
  qsstd(
    sim_unif[ , "percent_native_american_2010"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



st_fit <- sstdFit(data_for_var$percent_asian_2010)
sim$percent_asian_2010 <- 
  qsstd(
    sim_unif[ , "percent_asian_2010"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



sim$Percent_Diabetes_2013<-qnorm(sim_unif[,"Percent_Diabetes_2013"], mean= 11.25921 ,sd=2.460515)
sim$Population_Estimate_2013<-qnorm(sim_unif[,"Population_Estimate_2013"], mean=10.35833 , sd = 1.443346 )
sim$percent_65_and_older_2010<-qnorm(sim_unif[,"percent_65_and_older_2010"], mean=15.92134 ,sd=4.100516 )
sim$percent_18_or_younger_2010<-qnorm(sim_unif[,"percent_18_or_younger_2010"], mean=23.42614 ,sd=3.188879 )
sim$Median_household_income_2015<-qnorm(sim_unif[,"Median_household_income_2015"], mean=10.76431  ,sd=0.2315657  )
# a lot of liberties on this one 


st_fit <- sstdFit(data_for_var$Satewide_food_insecurity_12_14)
sim$Satewide_food_insecurity_12_14 <- 
  qsstd(
    sim_unif[ , "Satewide_food_insecurity_12_14"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

sim$Fast_Food_per_thousand_2011<-truncnorm::qtruncnorm(sim_unif[,"Fast_Food_per_thousand_2011"], a=0,mean=0.5, sd=0.27)
sim$percent_household_low_access_2015<-truncnorm::qtruncnorm(sim_unif[,"Fast_Food_per_thousand_2011"], a=0,b=100,mean=19, sd=18.25)

st_fit <- sstdFit(data_for_var$Poverty_rate_2015)
sim$Poverty_rate_2015 <- 
  qsstd(
    sim_unif[ , "Satewide_food_insecurity_12_14"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )


st_fit <- sstdFit(data_for_var$percent_hispanic_2010)
sim$percent_hispanic_2010 <- 
  qsstd(
    sim_unif[ , "percent_hispanic_2010"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



st_fit <- sstdFit(data_for_var$Full_service_restauraunt_per_thousand_2011)
sim$Full_service_restauraunt_per_thousand_2011 <- 
  qsstd(
    sim_unif[ , "Full_service_restauraunt_per_thousand_2011"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

st_fit <- sstdFit(data_for_var$Grocery_per_thousand_2011)
T_Grocery_per_thousand_2011 <- 
  qsstd(
    sim_unif[ , "Grocery_per_thousand_2011"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )


st_fit <- sstdFit(data_for_var$Farmers_markets_per_capita_2013)
sim$Farmers_markets_per_capita_2013 <- 
  qsstd(
    sim_unif[ , "Farmers_markets_per_capita_2013"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

st_fit <- sstdFit(data_for_var$Statewide_percent_obesity_2012)
sim$Statewide_percent_obesity_2012 <- 
  qsstd(
    sim_unif[ , "Statewide_percent_obesity_2012"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



st_fit <- sstdFit(data_for_var$Statewide_percent_obesity_2012)
sim$Statewide_percent_obesity_2012 <- 
  qsstd(
    sim_unif[ , "Statewide_percent_obesity_2012"], 
    mean = st_fit$estimate["mean"], 
    sd = 3.5, 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

st_fit <- sstdFit(data_for_var$Specialty_store_per_thousand_2011)
sim$Specialty_store_per_thousand_2011 <- 
  qsstd(
    sim_unif[ , "Specialty_store_per_thousand_2011"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )


# Observed probabilities:
probs1 <- table(data_for_var$FOOD_BANKS_2018) / nrow(data_for_var)
# Cumulative:
cum_probs1 <- cumsum(probs1)
# Add zero:
cum_probs1 <- c(0, cum_probs1)
# So we need to split the data:
cut(x = sim_unif[ ,"FOOD_BANKS_2018"], breaks = cum_probs1)
# Take it as the same integers from which we created the margin:
sim$FOOD_BANKS_2018 <- as.integer(cut(x = sim_unif[ , "FOOD_BANKS_2018"], breaks = cum_probs1))
# Check proportions:
table(sim$FOOD_BANKS_2018)


# Observed probabilities:
probs2 <- table(data_for_var$Persistant_poverty_counties_2010) / nrow(data_for_var)
# Cumulative:
cum_probs2 <- cumsum(probs2)
# Add zero:
cum_probs2 <- c(0, cum_probs2)
# So we need to split the data:
cut(x = sim_unif[ ,"Persistant_poverty_counties_2010"], breaks = cum_probs2)
# Take it as the same integers from which we created the margin:
sim$Persistant_poverty_counties_2010 <- as.integer(cut(x = sim_unif[ , "Persistant_poverty_counties_2010"], breaks = cum_probs2))

# Observed probabilities:
probs3 <- table(data_for_var$metronon_metro_county_2010) / nrow(data_for_var)
# Cumulative:
cum_probs3 <- cumsum(probs3)
# Add zero:
cum_probs3 <- c(0, cum_probs3)
# So we need to split the data:
cut(x = sim_unif[ ,"metronon_metro_county_2010"], breaks = cum_probs3)
# Take it as the same integers from which we created the margin:
sim$metronon_metro_county_2010 <- as.integer(cut(x = sim_unif[ , "metronon_metro_county_2010"], breaks = cum_probs3))
# Check proportions:

# Observed probabilities:
probs4 <- table(data_for_var$population_loss_county_2010) / nrow(data_for_var)
# Cumulative:
cum_probs4 <- cumsum(probs4)
# Add zero:
cum_probs4 <- c(0, cum_probs4)
# So we need to split the data:
cut(x = sim_unif[ ,"population_loss_county_2010"], breaks = cum_probs4)
# Take it as the same integers from which we created the margin:
sim$population_loss_county_2010 <- as.integer(cut(x = sim_unif[ , "population_loss_county_2010"], breaks = cum_probs4))
# Check proportions:
table(sim$population_loss_county_2010)
table(data_for_var$population_loss_county_2010)
sim$population_loss_county_2010 
identical(sim$population_loss_county_2010,data_for_var$population_loss_county_2010)

#Observed probabilities:
probs5 <- table(data_for_var$region) / nrow(data_for_var)
# Cumulative:
cum_probs5 <- cumsum(probs5)
# Add zero:
cum_probs5 <- c(0, cum_probs5)
# So we need to split the data:
cut(x = sim_unif[ ,"region"], breaks = cum_probs5)
# Take it as the same integers from which we created the margin:
sim$region <- as.integer(cut(x = sim_unif[ , "region"], breaks = cum_probs5))
# Check proportions:
table(sim$region)
table(data_for_var$region)


#Observed probabilities:
probs6 <- table(data_for_var$highlowDM) / nrow(data_for_var)
# Cumulative:
cum_probs6 <- cumsum(probs6)
# Add zero:
cum_probs6 <- c(0, cum_probs6)
# So we need to split the data:
cut(x = sim_unif[ ,"highlowDM"], breaks = cum_probs6)
# Take it as the same integers from which we created the margin:
sim$highlowDM <- as.integer(cut(x = sim_unif[ , "highlowDM"], breaks = cum_probs6))
# Check proportions:
table(sim$highlowDM)
table(data_for_var$highlowDM)



#dropping the variables I couldnt fit 
drop<-c("Recreation_and_fitness_facilties_per_1000_2011","Percent_farmers_market_accepting_SNAP_2013", "Statewide_SNAP_participation_rate_2011","Supercenter_per_thousand_2011")
sim_variables<-sim[,!(names(sim) %in% drop)]


#setting"TEST" linear model
sim_regression_test<-sim_variables[ ,-30]
test_matrix<-model.matrix(Percent_Diabetes_2013~.,sim_regression_test,nvmax=28)
true_model_test<-lm(sim_regression_test$Percent_Diabetes_2013~., data=sim_regression_test)
true_F_X<-predict(true_model_test, data=sim_regression_test)
```{r}
# test set


set.seed(2)
sim_mat <- mvtnorm::rmvnorm(nrow(data_for_var), sigma = cor(data_for_var))
sim_unif <- pnorm(sim_mat)
colnames(sim_unif)<-colnames(data_for_var)


sim<-data_for_var

sim$percent_black_2010 <- qgamma(sim_unif[,"percent_black_2010"], shape=0.4402158, rate = 0.05254412 )
sim$Direct_farm_sales_per_capita_2012 <- qgamma(sim_unif[,"Direct_farm_sales_per_capita_2012"], shape=0.5355279 , rate = 0.06477432 )
sim$percent_white_2010<-truncnorm::qtruncnorm(sim_unif[,"percent_white_2010"], a=0, b=100, mean=160, sd = 45)
sim$percent_pacific_islander_2010<-qgamma(sim_unif[,"percent_pacific_islander_2010"], shape=0.2740278  , rate = 4.224056)
sim$CSA_2013<-qgamma(sim_unif[,"CSA_2013"], shape=0.3002073  , rate = 2465.577  )

st_fit <- sstdFit(data_for_var$percent_native_american_2010)
sim$percent_native_american_2010 <- 
  qsstd(
    sim_unif[ , "percent_native_american_2010"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



st_fit <- sstdFit(data_for_var$percent_asian_2010)
sim$percent_asian_2010 <- 
  qsstd(
    sim_unif[ , "percent_asian_2010"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



sim$Percent_Diabetes_2013<-qnorm(sim_unif[,"Percent_Diabetes_2013"], mean= 11.25921 ,sd=2.460515)
sim$Population_Estimate_2013<-qnorm(sim_unif[,"Population_Estimate_2013"], mean=10.35833 , sd = 1.443346 )
sim$percent_65_and_older_2010<-qnorm(sim_unif[,"percent_65_and_older_2010"], mean=15.92134 ,sd=4.100516 )
sim$percent_18_or_younger_2010<-qnorm(sim_unif[,"percent_18_or_younger_2010"], mean=23.42614 ,sd=3.188879 )
sim$Median_household_income_2015<-qnorm(sim_unif[,"Median_household_income_2015"], mean=10.76431  ,sd=0.2315657  )
# a lot of liberties on this one 


st_fit <- sstdFit(data_for_var$Satewide_food_insecurity_12_14)
sim$Satewide_food_insecurity_12_14 <- 
  qsstd(
    sim_unif[ , "Satewide_food_insecurity_12_14"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

sim$Fast_Food_per_thousand_2011<-truncnorm::qtruncnorm(sim_unif[,"Fast_Food_per_thousand_2011"], a=0,mean=0.5, sd=0.27)
sim$percent_household_low_access_2015<-truncnorm::qtruncnorm(sim_unif[,"Fast_Food_per_thousand_2011"], a=0,b=100,mean=19, sd=18.25)

st_fit <- sstdFit(data_for_var$Poverty_rate_2015)
sim$Poverty_rate_2015 <- 
  qsstd(
    sim_unif[ , "Satewide_food_insecurity_12_14"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )


st_fit <- sstdFit(data_for_var$percent_hispanic_2010)
sim$percent_hispanic_2010 <- 
  qsstd(
    sim_unif[ , "percent_hispanic_2010"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



st_fit <- sstdFit(data_for_var$Full_service_restauraunt_per_thousand_2011)
sim$Full_service_restauraunt_per_thousand_2011 <- 
  qsstd(
    sim_unif[ , "Full_service_restauraunt_per_thousand_2011"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

st_fit <- sstdFit(data_for_var$Grocery_per_thousand_2011)
T_Grocery_per_thousand_2011 <- 
  qsstd(
    sim_unif[ , "Grocery_per_thousand_2011"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )


st_fit <- sstdFit(data_for_var$Farmers_markets_per_capita_2013)
sim$Farmers_markets_per_capita_2013 <- 
  qsstd(
    sim_unif[ , "Farmers_markets_per_capita_2013"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

st_fit <- sstdFit(data_for_var$Statewide_percent_obesity_2012)
sim$Statewide_percent_obesity_2012 <- 
  qsstd(
    sim_unif[ , "Statewide_percent_obesity_2012"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



st_fit <- sstdFit(data_for_var$Statewide_percent_obesity_2012)
sim$Statewide_percent_obesity_2012 <- 
  qsstd(
    sim_unif[ , "Statewide_percent_obesity_2012"], 
    mean = st_fit$estimate["mean"], 
    sd = 3.5, 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

st_fit <- sstdFit(data_for_var$Specialty_store_per_thousand_2011)
sim$Specialty_store_per_thousand_2011 <- 
  qsstd(
    sim_unif[ , "Specialty_store_per_thousand_2011"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )


# Observed probabilities:
probs1 <- table(data_for_var$FOOD_BANKS_2018) / nrow(data_for_var)
# Cumulative:
cum_probs1 <- cumsum(probs1)
# Add zero:
cum_probs1 <- c(0, cum_probs1)
# So we need to split the data:
cut(x = sim_unif[ ,"FOOD_BANKS_2018"], breaks = cum_probs1)
# Take it as the same integers from which we created the margin:
sim$FOOD_BANKS_2018 <- as.integer(cut(x = sim_unif[ , "FOOD_BANKS_2018"], breaks = cum_probs1))
# Check proportions:
table(sim$FOOD_BANKS_2018)


# Observed probabilities:
probs2 <- table(data_for_var$Persistant_poverty_counties_2010) / nrow(data_for_var)
# Cumulative:
cum_probs2 <- cumsum(probs2)
# Add zero:
cum_probs2 <- c(0, cum_probs2)
# So we need to split the data:
cut(x = sim_unif[ ,"Persistant_poverty_counties_2010"], breaks = cum_probs2)
# Take it as the same integers from which we created the margin:
sim$Persistant_poverty_counties_2010 <- as.integer(cut(x = sim_unif[ , "Persistant_poverty_counties_2010"], breaks = cum_probs2))

# Observed probabilities:
probs3 <- table(data_for_var$metronon_metro_county_2010) / nrow(data_for_var)
# Cumulative:
cum_probs3 <- cumsum(probs3)
# Add zero:
cum_probs3 <- c(0, cum_probs3)
# So we need to split the data:
cut(x = sim_unif[ ,"metronon_metro_county_2010"], breaks = cum_probs3)
# Take it as the same integers from which we created the margin:
sim$metronon_metro_county_2010 <- as.integer(cut(x = sim_unif[ , "metronon_metro_county_2010"], breaks = cum_probs3))
# Check proportions:

# Observed probabilities:
probs4 <- table(data_for_var$population_loss_county_2010) / nrow(data_for_var)
# Cumulative:
cum_probs4 <- cumsum(probs4)
# Add zero:
cum_probs4 <- c(0, cum_probs4)
# So we need to split the data:
cut(x = sim_unif[ ,"population_loss_county_2010"], breaks = cum_probs4)
# Take it as the same integers from which we created the margin:
sim$population_loss_county_2010 <- as.integer(cut(x = sim_unif[ , "population_loss_county_2010"], breaks = cum_probs4))
# Check proportions:
table(sim$population_loss_county_2010)
table(data_for_var$population_loss_county_2010)
sim$population_loss_county_2010 
identical(sim$population_loss_county_2010,data_for_var$population_loss_county_2010)

#Observed probabilities:
probs5 <- table(data_for_var$region) / nrow(data_for_var)
# Cumulative:
cum_probs5 <- cumsum(probs5)
# Add zero:
cum_probs5 <- c(0, cum_probs5)
# So we need to split the data:
cut(x = sim_unif[ ,"region"], breaks = cum_probs5)
# Take it as the same integers from which we created the margin:
sim$region <- as.integer(cut(x = sim_unif[ , "region"], breaks = cum_probs5))
# Check proportions:
table(sim$region)
table(data_for_var$region)


#Observed probabilities:
probs6 <- table(data_for_var$highlowDM) / nrow(data_for_var)
# Cumulative:
cum_probs6 <- cumsum(probs6)
# Add zero:
cum_probs6 <- c(0, cum_probs6)
# So we need to split the data:
cut(x = sim_unif[ ,"highlowDM"], breaks = cum_probs6)
# Take it as the same integers from which we created the margin:
sim$highlowDM <- as.integer(cut(x = sim_unif[ , "highlowDM"], breaks = cum_probs6))
# Check proportions:
table(sim$highlowDM)
table(data_for_var$highlowDM)



#dropping the variables I couldnt fit 
drop<-c("Recreation_and_fitness_facilties_per_1000_2011","Percent_farmers_market_accepting_SNAP_2013", "Statewide_SNAP_participation_rate_2011","Supercenter_per_thousand_2011")
sim_variables<-sim[,!(names(sim) %in% drop)]


#setting"TEST" linear model
sim_regression_test<-sim_variables[ ,-30]
test_matrix<-model.matrix(Percent_Diabetes_2013~.,sim_regression_test,nvmax=28)
true_model_test<-lm(sim_regression_test$Percent_Diabetes_2013~., data=sim_regression_test)
true_F_X<-predict(true_model_test, data=sim_regression_test)
```



```{r}
library(leaps)
library(tidyverse)

pred_test<-NULL
predictions <- NULL
M<-100
df_seq<-seq(1:28)
for (m in 1:M)
{
sim_mat <- mvtnorm::rmvnorm(nrow(data_for_var), sigma = cor(data_for_var))
sim_unif <- pnorm(sim_mat)
colnames(sim_unif)<-colnames(data_for_var)


sim<-data_for_var

sim$percent_black_2010 <- qgamma(sim_unif[,"percent_black_2010"], shape=0.4402158, rate = 0.05254412 )
sim$Direct_farm_sales_per_capita_2012 <- qgamma(sim_unif[,"Direct_farm_sales_per_capita_2012"], shape=0.5355279 , rate = 0.06477432 )
sim$percent_white_2010<-truncnorm::qtruncnorm(sim_unif[,"percent_white_2010"], a=0, b=100, mean=160, sd = 45)
sim$percent_pacific_islander_2010<-qgamma(sim_unif[,"percent_pacific_islander_2010"], shape=0.2740278  , rate = 4.224056)
sim$CSA_2013<-qgamma(sim_unif[,"CSA_2013"], shape=0.3002073  , rate = 2465.577  )

st_fit <- sstdFit(data_for_var$percent_native_american_2010)
sim$percent_native_american_2010 <- 
  qsstd(
    sim_unif[ , "percent_native_american_2010"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



st_fit <- sstdFit(data_for_var$percent_asian_2010)
sim$percent_asian_2010 <- 
  qsstd(
    sim_unif[ , "percent_asian_2010"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



sim$Percent_Diabetes_2013<-qnorm(sim_unif[,"Percent_Diabetes_2013"], mean= 11.25921 ,sd=2.460515)
sim$Population_Estimate_2013<-qnorm(sim_unif[,"Population_Estimate_2013"], mean=10.35833 , sd = 1.443346 )
sim$percent_65_and_older_2010<-qnorm(sim_unif[,"percent_65_and_older_2010"], mean=15.92134 ,sd=4.100516 )
sim$percent_18_or_younger_2010<-qnorm(sim_unif[,"percent_18_or_younger_2010"], mean=23.42614 ,sd=3.188879 )
sim$Median_household_income_2015<-qnorm(sim_unif[,"Median_household_income_2015"], mean=10.76431  ,sd=0.2315657  )
# a lot of liberties on this one 


st_fit <- sstdFit(data_for_var$Satewide_food_insecurity_12_14)
sim$Satewide_food_insecurity_12_14 <- 
  qsstd(
    sim_unif[ , "Satewide_food_insecurity_12_14"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

sim$Fast_Food_per_thousand_2011<-truncnorm::qtruncnorm(sim_unif[,"Fast_Food_per_thousand_2011"], a=0,mean=0.5, sd=0.27)
sim$percent_household_low_access_2015<-truncnorm::qtruncnorm(sim_unif[,"Fast_Food_per_thousand_2011"], a=0,b=100,mean=19, sd=18.25)

st_fit <- sstdFit(data_for_var$Poverty_rate_2015)
sim$Poverty_rate_2015 <- 
  qsstd(
    sim_unif[ , "Satewide_food_insecurity_12_14"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )


st_fit <- sstdFit(data_for_var$percent_hispanic_2010)
sim$percent_hispanic_2010 <- 
  qsstd(
    sim_unif[ , "percent_hispanic_2010"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



st_fit <- sstdFit(data_for_var$Full_service_restauraunt_per_thousand_2011)
sim$Full_service_restauraunt_per_thousand_2011 <- 
  qsstd(
    sim_unif[ , "Full_service_restauraunt_per_thousand_2011"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

st_fit <- sstdFit(data_for_var$Grocery_per_thousand_2011)
T_Grocery_per_thousand_2011 <- 
  qsstd(
    sim_unif[ , "Grocery_per_thousand_2011"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )


st_fit <- sstdFit(data_for_var$Farmers_markets_per_capita_2013)
sim$Farmers_markets_per_capita_2013 <- 
  qsstd(
    sim_unif[ , "Farmers_markets_per_capita_2013"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

st_fit <- sstdFit(data_for_var$Statewide_percent_obesity_2012)
sim$Statewide_percent_obesity_2012 <- 
  qsstd(
    sim_unif[ , "Statewide_percent_obesity_2012"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )



st_fit <- sstdFit(data_for_var$Statewide_percent_obesity_2012)
sim$Statewide_percent_obesity_2012 <- 
  qsstd(
    sim_unif[ , "Statewide_percent_obesity_2012"], 
    mean = st_fit$estimate["mean"], 
    sd = 3.5, 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )

st_fit <- sstdFit(data_for_var$Specialty_store_per_thousand_2011)
sim$Specialty_store_per_thousand_2011 <- 
  qsstd(
    sim_unif[ , "Specialty_store_per_thousand_2011"], 
    mean = st_fit$estimate["mean"], 
    sd = st_fit$estimate["sd"], 
    xi = st_fit$estimate["xi"], 
    nu = st_fit$estimate["nu"]
    )


# Observed probabilities:
probs1 <- table(data_for_var$FOOD_BANKS_2018) / nrow(data_for_var)
# Cumulative:
cum_probs1 <- cumsum(probs1)
# Add zero:
cum_probs1 <- c(0, cum_probs1)
# So we need to split the data:
cut(x = sim_unif[ ,"FOOD_BANKS_2018"], breaks = cum_probs1)
# Take it as the same integers from which we created the margin:
sim$FOOD_BANKS_2018 <- as.integer(cut(x = sim_unif[ , "FOOD_BANKS_2018"], breaks = cum_probs1))
# Check proportions:
table(sim$FOOD_BANKS_2018)


# Observed probabilities:
probs2 <- table(data_for_var$Persistant_poverty_counties_2010) / nrow(data_for_var)
# Cumulative:
cum_probs2 <- cumsum(probs2)
# Add zero:
cum_probs2 <- c(0, cum_probs2)
# So we need to split the data:
cut(x = sim_unif[ ,"Persistant_poverty_counties_2010"], breaks = cum_probs2)
# Take it as the same integers from which we created the margin:
sim$Persistant_poverty_counties_2010 <- as.integer(cut(x = sim_unif[ , "Persistant_poverty_counties_2010"], breaks = cum_probs2))

# Observed probabilities:
probs3 <- table(data_for_var$metronon_metro_county_2010) / nrow(data_for_var)
# Cumulative:
cum_probs3 <- cumsum(probs3)
# Add zero:
cum_probs3 <- c(0, cum_probs3)
# So we need to split the data:
cut(x = sim_unif[ ,"metronon_metro_county_2010"], breaks = cum_probs3)
# Take it as the same integers from which we created the margin:
sim$metronon_metro_county_2010 <- as.integer(cut(x = sim_unif[ , "metronon_metro_county_2010"], breaks = cum_probs3))
# Check proportions:

# Observed probabilities:
probs4 <- table(data_for_var$population_loss_county_2010) / nrow(data_for_var)
# Cumulative:
cum_probs4 <- cumsum(probs4)
# Add zero:
cum_probs4 <- c(0, cum_probs4)
# So we need to split the data:
cut(x = sim_unif[ ,"population_loss_county_2010"], breaks = cum_probs4)
# Take it as the same integers from which we created the margin:
sim$population_loss_county_2010 <- as.integer(cut(x = sim_unif[ , "population_loss_county_2010"], breaks = cum_probs4))
# Check proportions:
table(sim$population_loss_county_2010)
table(data_for_var$population_loss_county_2010)
sim$population_loss_county_2010 
identical(sim$population_loss_county_2010,data_for_var$population_loss_county_2010)

#Observed probabilities:
probs5 <- table(data_for_var$region) / nrow(data_for_var)
# Cumulative:
cum_probs5 <- cumsum(probs5)
# Add zero:
cum_probs5 <- c(0, cum_probs5)
# So we need to split the data:
cut(x = sim_unif[ ,"region"], breaks = cum_probs5)
# Take it as the same integers from which we created the margin:
sim$region <- as.integer(cut(x = sim_unif[ , "region"], breaks = cum_probs5))
# Check proportions:
table(sim$region)
table(data_for_var$region)


#Observed probabilities:
probs6 <- table(data_for_var$highlowDM) / nrow(data_for_var)
# Cumulative:
cum_probs6 <- cumsum(probs6)
# Add zero:
cum_probs6 <- c(0, cum_probs6)
# So we need to split the data:
cut(x = sim_unif[ ,"highlowDM"], breaks = cum_probs6)
# Take it as the same integers from which we created the margin:
sim$highlowDM <- as.integer(cut(x = sim_unif[ , "highlowDM"], breaks = cum_probs6))
# Check proportions:
table(sim$highlowDM)
table(data_for_var$highlowDM)



#dropping the variables I couldnt fit 
drop<-c("Recreation_and_fitness_facilties_per_1000_2011","Percent_farmers_market_accepting_SNAP_2013", "Statewide_SNAP_participation_rate_2011","Supercenter_per_thousand_2011")
sim_variables<-sim[,!(names(sim) %in% drop)]


sim_regression_train<-sim_variables[ ,-30]
train_matrix<-model.matrix(Percent_Diabetes_2013~.,sim_regression_train,nvmax=28)
regression_best <- regsubsets(Percent_Diabetes_2013 ~., sim_regression_train,nvmax =28, method = "forward")
 

for (i in 1:length(df_seq)) {
    
   coef <- coef(regression_best, id = i)
    pred_test$pred_f_x <- (test_matrix[ ,names(coef)] %*% coef)
    pred_test$true_f_x = true_F_X
    pred_test$true_y = sim_regression_test$Percent_Diabetes_2013
    pred_test$test_obs_idx = seq(1:2883)
    pred_test$m = m
    pred_test$df= df_seq[i]
    predictions <- bind_rows(predictions, pred_test)
  }



}

bias_var_MSE <-
  predictions %>%
  group_by(test_obs_idx, df) %>%
  summarize(
    pt_wise_bias_sq = (mean(pred_f_x - true_f_x)) ^ 2,
    pt_wise_var = mean( (pred_f_x - mean(pred_f_x)) ^ 2 ),
    pt_wise_MSEE = mean( (pred_f_x - true_f_x) ^ 2 ),
    pt_wise_MSEP = mean( (pred_f_x - true_y) ^ 2 )
  ) %>%
  group_by(df) %>%
  summarize(
    bias_sq = mean(pt_wise_bias_sq),
    var = mean(pt_wise_var),
    MSEE = mean(pt_wise_MSEE),
    MSEP = mean(pt_wise_MSEP)
  )

(g <-
  bias_var_MSE %>%
  pivot_longer(c(bias_sq, var, MSEE, MSEP)) %>% 
  ggplot() +
  geom_line(aes(x = df, y = value, color = name)) +
  # geom_hline(yintercept = est_var_epsilon, lty = 2) +
  theme_bw() +
  xlab("Flexibility") +
  ylab("Mean Squared Error") +
  theme(legend.title = element_blank()))


```
