---
title: "hw7_duc_submission"
author: "Stephen Shannon"
date: "2025-04-16"
output: html_document
---


```{r}
library(readxl)
library(dplyr)

access <- read_excel("C:/1. Học tập/Spring 2025/4. Stat 630/Final project/FoodEnvironmentAtlas (After converted).xlsx", sheet = "ACCESS")
health <- read_excel("C:/1. Học tập/Spring 2025/4. Stat 630/Final project/FoodEnvironmentAtlas (After converted).xlsx", sheet = "HEALTH")
socioeconomic <- read_excel("C:/1. Học tập/Spring 2025/4. Stat 630/Final project/FoodEnvironmentAtlas (After converted).xlsx", sheet = "SOCIOECONOMIC")

data <- access %>%
  inner_join(health, by = c("FIPS", "State", "County")) %>%
  inner_join(socioeconomic, by = c("FIPS", "State", "County"))

head(data)
```

### Task 1: Exploratory Data Analysis (Pairwise plots)

```{r}

# install.packages("GGally")
library(GGally)

num_vars <- data %>%
  select_if(is.numeric)

num_vars_subset <- num_vars[, 1:10]

ggpairs(num_vars_subset)

```

### Task 2: Quantitative outcome analysis

**Part a: Marginal simple linear regression**

```{r}
lm_simple <- lm(PCT_OBESE_ADULTS17 ~ PCT_LACCESS_POP15, data = data)
summary(lm_simple)
```

**Part b: Multiple linear regression with polynomial and interaction terms**

```{r}
lm_multiple <- lm(PCT_OBESE_ADULTS17 ~ PCT_LACCESS_POP15 + I(PCT_LACCESS_POP15^2) + 
                  PCT_DIABETES_ADULTS13 + PCT_NHBLACK10 + 
                  PCT_LACCESS_POP15:PCT_DIABETES_ADULTS13 + 
                  PCT_LACCESS_POP15:PCT_NHBLACK10,
                  data = data)

summary(lm_multiple)
```

### Task 3: Qualitative outcome analysis

**Part a: KNN (K-nearest neighbors)**

```{r}

# install.packages("caret")
library(dplyr)
library(caret)

data_knn <- data %>%
  dplyr::select(METRO13, PCT_OBESE_ADULTS17, PCT_DIABETES_ADULTS13, PCT_NHBLACK10) %>%
  na.omit()

set.seed(123)
trainIndex <- createDataPartition(data_knn$METRO13, p = 0.9, list = FALSE)
train_knn <- data_knn[trainIndex,]
test_knn <- data_knn[-trainIndex,]

knn_fit <- train(as.factor(METRO13) ~ ., data = train_knn, method = "knn", tuneLength = 5)
print(knn_fit)
```

**Part b: Logistic regression with polynomial and interaction terms**

```{r}
logistic_model <- glm(METRO13 ~ PCT_OBESE_ADULTS17 + I(PCT_OBESE_ADULTS17^2) + 
                      PCT_DIABETES_ADULTS13 + 
                      PCT_OBESE_ADULTS17:PCT_DIABETES_ADULTS13 +
                      PCT_OBESE_ADULTS17:PCT_NHBLACK10,
                      data = train_knn, family = binomial)

summary(logistic_model)
```

**Part c: Linear discriminant analysis (LDA)**

```{r}

library(MASS)

lda_fit <- lda(as.factor(METRO13) ~ ., data = train_knn)
lda_pred <- predict(lda_fit, test_knn)
table(lda_pred$class, test_knn$METRO13)
```

**Part d: Quadratic discriminant analysis (QDA)**

```{r}

qda_fit <- qda(as.factor(METRO13) ~ ., data = train_knn)
qda_pred <- predict(qda_fit, test_knn)
table(qda_pred$class, test_knn$METRO13)
```

**Part e: Naive Bayes (two kernels: Gaussian & Kernel density)**

```{r}

library(e1071)

nb_gaussian <- naiveBayes(as.factor(METRO13) ~ ., data = train_knn)
nb_gaussian_pred <- predict(nb_gaussian, test_knn)
table(nb_gaussian_pred, test_knn$METRO13)

nb_kernel <- train(as.factor(METRO13) ~ ., data = train_knn, method = "nb",
                   trControl=trainControl(method="cv", number=10))
print(nb_kernel)

```